---
title: "Assignment_2"
author: "Joseph Demore"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Small tidbits about my style;
```

=== logically equivalent to;
== is equal to
{ } usaully used to point to a variable name or choice that varies from directions
        (i try to make my life easier from a coding side)

```
        
```{r}
# short block to check if the packages are installed; will download and install any that are not.
packages <- c("class", "caret", "gmodels")

for (pkg in packages) {
  if (!requireNamespace(pkg, quietly = TRUE)) install.packages(pkg, repos = "https://cloud.r-project.org")
  library(pkg, character.only = TRUE)
}
```

### Helpers / things I reused.

```{r}
# helper fn's

# fn to count unique
uCount <- function(x) {
  length(unique(x))
}

# normalizations for data
normalize <- function(data, model) {
  data[qVars] <- predict(model, data[qVars])
  data[predictors]
}

# accuracy fn for knn
knnAcc <- function(k, trainX, testX, trainY, testY) {
  mean(knn(trainX, testX, cl = trainY, k = k) == testY)
}

# cross table helper
knnCross <- function(k, trainX, testX, trainY, testY) {
  CrossTable(testY, knn(trainX, testX, cl = trainY, k = k), prop.chisq = FALSE)
}

# summary for the knn { used with end/knnCross comparison }
knnSummary <- function(k, trainX, testX, trainY, testY) {
  cm <- table(factor(testY, levels = c("0", "1")),
              factor(knn(trainX, testX, cl = trainY, k = k), levels = c("0", "1")))
  TN <- cm[1, 1]; FP <- cm[1, 2]; FN <- cm[2, 1]; TP <- cm[2, 2]
  data.frame(n = sum(cm), TN, FP, FN, TP,
             Accuracy = (TP + TN) / sum(cm),
             Sensitivity = TP / (TP + FN),
             Specificity = TN / (TN + FP),
             Precision = TP / (TP + FP))
}
```

### Data Setup & Inspection
Sorted Ascending by unique values to help id cat's. not dogs. cats.

```{r}
# Reads UniversalBank.csv; str shows structure (colnam, type, 10 values)
uBank <- read.csv("UniversalBank.csv")
str(uBank)

# summary to check data
summary(uBank)

# uCount on each col to help identify categoricals
sapply(uBank, uCount)

# sorted
sort(sapply(uBank, uCount))

# categoricals: 
# Personal.Loan [target outcome], Securities.Account, CD.Account, Online,
# CreditCard, Education [this needs to be split]
# Zip Code and ID (arith not going to do anything here) but we're dropping

# Numeric:
# Family (even though count is small), Age, Experience, CCAvg, Income, Mortgage

```
### Bad data || but instructions say to keep all predictors except-listed.
It was seen.
Decision to keep was made { assignment asked for predictors to be kept }.
Results would have likely been different without the 52 potentially bad-data driven FNs.

```{r}
# summary showed -3 experience value;... assuming this is correlated to -3 years
# it's likely bad data -- these rows //should// be removed.
# especially since all - experience customers declined the loan.
summary(uBank$Experience)
head(sort(uBank$Experience))
head(uBank[uBank$Experience < 0, ])
nrow(uBank[uBank$Experience < 0, c("Experience", "Personal.Loan")])
# verify they're all negative. true == all declined.
# this is === if ALL people with negative experience are declined, print true
all(uBank$Personal.Loan[uBank$Experience<0] == 0)
```


### Customer definition from the assignment (for q1, q4)

```{r}
sCust <- data.frame(
  Age = 40,
  Experience = 10,
  Income = 84,
  Family = 2,
  CCAvg = 2,
  Education_1 = 0,
  Education_2 = 1,
  Education_3 = 0,
  Mortgage = 0,
  Securities.Account = 0,
  CD.Account = 0,
  Online = 1,
  CreditCard = 1
)
```

### Sample has Edu1,Edu2, Edu3 so we can factor edu
Set diff shows id / zip / edu / pers loan as differences
once edu is factored we will remove id, zip, personal loan
(though i would have assumed zip would be relevant, location matters).
It would have been cool to geoplot them on a heatmap.

```{r}
setdiff(names(uBank), names(sCust))
```
### Drop ID, Zip then split nonbinary categoricals

```{r}
# categorizes education via factor, binds to df and removes dropped cols
# note sep by _ is only there to match the sample customer.
uBank$Education <- factor(uBank$Education)
uBank <- cbind(uBank, predict(dummyVars(~ Education, data = uBank, sep = "_"), uBank))
uBank$Education <- NULL
uBank$ID <- NULL
uBank$ZIP.Code <- NULL

summary(uBank)
```

```{r}
# only remaining Personal.Loan (acceptance) is missing from the sample customer.
setdiff(names(uBank), names(sCust))
```
### Let's do some ML.. finally.

```{r}
# this one needs to be categorical since it's our target. other cat binarys need
# to stay 0-1 for euclid distance on knn.
uBank$Personal.Loan <- factor(uBank$Personal.Loan)

# make our randomness not randomness at all.
set.seed(1234567890)
# partition for t&v
split = createDataPartition(uBank$Personal.Loan, p=0.6, list = FALSE)
# training gets 60% and validation 40% // split indexes at 60% of data
training <- uBank[split,]
validation <- uBank[-split,]
# define predictors; which are all cols except personal loan //via setdiff
predictors <- setdiff(names(uBank), "Personal.Loan")
# define quant & cat: setdiff predictors&qVars to get cats. === (1- X) for names
qVars <- c("Age", "Experience", "Income", "Family", "CCAvg", "Mortgage")
cVars <- setdiff(predictors, qVars)
```

### Normalization...
Note I defined this above myself.

```{r}
# normalization of model, training, validation, and sample customer
normModel <- preProcess(training[qVars])
tNorm <- normalize(training, normModel)
vNorm <- normalize(validation, normModel)
cNorm <- normalize(sCust, normModel)
```

# **Question 1** k-NN for sCust
Success as 1 (loan accept) && cutoff 0.5.
-- with this one; cutoff is irrelevant since nearest neighbor will be 0 or 1

```{r}
knn(tNorm, cNorm, cl = training$Personal.Loan, k=1)
```
### Classified as not accepted [Level 0].

# **Question 2** Choice of k that balances over-fitting&ignorance

```{r}
# odds here >> evens produce ties between yes-no
kTest <- seq(1, 49, by = 2)
acc <- sapply(kTest, knnAcc, tNorm, vNorm, training$Personal.Loan, validation$Personal.Loan)
acc

# exclude k==1 since it would just mimic its closest neighbor & that's
# the definition of overfitting.
begone <- kTest == 1
choice <- kTest[!begone][which.max(acc[!begone])]
choice
max(acc[!begone])
plot(kTest, acc)
```

### I would choose k==5 for this, as it has the highest accuracy for k != 1.
Plot for great justice. Trend shows consistent downward tendency.
k==49, the accuracy falls to ~92.6%, which is pretty close to the naive baseline.
Larger K would ignore the predictors here.

# **Question 3** Confusion Matrix for validation on best  k { choice } var
Cross table shows Acc 95.7%(1806+108 TN,TP/Total), but missed more than 40% of acceptance {sens==56%}.
Still concerned about the negative experience data rows in our sample.

```{r}
# Cross table {def above} 
knnCross(choice, tNorm, vNorm, training$Personal.Loan, validation$Personal.Loan)

# summary fn ^^above used here
q3Summary <- rbind(
  Validation = knnSummary(choice, tNorm, vNorm, training$Personal.Loan, validation$Personal.Loan)
)
round(q3Summary, 4)
```

# **Question 4** Classify sCust with best k { choice } of 5

```{r}
knn(tNorm, cNorm, cl = training$Personal.Loan, k = choice)
```

### Customer still gets rejected. He's having a bad day, we re-ran his credit.

# **Question 5** Repartition; training->validation->test. k-NN with 5 && cMatr
so.. do it all again. 

```{r}
# Three way split 50/30/20 // train first, then split the leftover
# reset seed so this split lands the same every run
set.seed(1234567890)

# training gets 50% (2500 rows)
q5Split <- createDataPartition(uBank$Personal.Loan, p = 0.5, list = FALSE)
q5Training <- uBank[q5Split, ]
# the other 50% is leftover (2500 rows)
q5Rest <- uBank[-q5Split, ]

# .6 of the leftover is validation (1500 = 30% of total); remaining is test (1000 = 20%)
q5RestSplit <- createDataPartition(q5Rest$Personal.Loan, p = 0.6, list = FALSE)
q5Validation <- q5Rest[q5RestSplit, ]
q5Test <- q5Rest[-q5RestSplit, ]

# normalize: fit on q5 training only, apply to all three
q5NormModel <- preProcess(q5Training[qVars])
q5tNorm <- normalize(q5Training, q5NormModel)
q5vNorm <- normalize(q5Validation, q5NormModel)
q5teNorm <- normalize(q5Test, q5NormModel)

# confusion matrices at best k { choice }: training on itself, then validation, then test
knnCross(choice, q5tNorm, q5tNorm, q5Training$Personal.Loan, q5Training$Personal.Loan)
knnCross(choice, q5tNorm, q5vNorm, q5Training$Personal.Loan, q5Validation$Personal.Loan)
knnCross(choice, q5tNorm, q5teNorm, q5Training$Personal.Loan, q5Test$Personal.Loan)

# summary table { built from fn ^^ above / near top } 
# training on itself, then validation, then test
# -- based on metrics in slides. recall===sensitivity
q5Summary <- rbind(
  Training = knnSummary(choice, q5tNorm, q5tNorm, q5Training$Personal.Loan, q5Training$Personal.Loan),
  Validation = knnSummary(choice, q5tNorm, q5vNorm, q5Training$Personal.Loan, q5Validation$Personal.Loan),
  Test = knnSummary(choice, q5tNorm, q5teNorm, q5Training$Personal.Loan, q5Test$Personal.Loan)
)
round(q5Summary, 4) 

```
### **Question 5** Comparison || k==5

```
From chart above - Accuracy for Training(97.48%), Validation(96.07%), and Test(96.10%)
Accuracy is the highest on the training set, because each row counts as 1 of the 5 knn's
Validation/Test sets don't carry this property, so the accuracy drops slightly.

Sensitivity{recall} shows much different values. Training(73.75%), Validation(59.72%) and Test(60.42%)
Validation and test are essentially equal. Training is higher because each training 
row counts itself as one of its 5 neighbors.

Specificity, Precision are both near-perfect for all cases. Very few FP in any set.

Most of the reported errors (excluding the exceptionally few FP, 2 in total) are from FN

Validation/Testing sets both are better indicators of actual model performance here,
since they're not counting a row as its own nearest neighbor. 

Sampling noise is likely the cause for the light differences in values of validation/testing.

False negatives would likely be from the low acceptance rate (9.6%). The nearest neighbors
of these small sets have a smaller chance of also having acceptance==1. Since the cutoff
is 0.5, this would mean 3 of 5 would have to accept; and without the self row count in training
it becomes less likely.

Overall - the models do not seem to be overfit, and there are very few false positives -
however, it only carries ~60% accuracy on a true positive {===sensitivity}. this value
could use more tuning to make the model better, but other metrics look good.
```
