---
title: "Assignment_2"
author: "Joseph Demore"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# short block to check if the packages are installed; will download and install any that are not.
packages <- c("class", "caret", "gmodels")

for (pkg in packages) {
  if (!requireNamespace(pkg, quietly = TRUE)) install.packages(pkg, repos = "https://cloud.r-project.org")
  library(pkg, character.only = TRUE)
}
```

```{r}
# helper fn's

# fn to count unique
uCount <- function(x) {
  length(unique(x))
}

```

### Data Setup & Inspection

Sorted Ascending by unique values; we can 

```{r}
# Reads UniversalBank.csv; str shows structure (colnam, type, 10 values)
uBank <- read.csv("UniversalBank.csv")
str(uBank)

# summary to check data
summary(uBank)

# uCount on each col to help identify categoricals
sapply(uBank, uCount)

# sorted
sort(sapply(uBank, uCount))

# categoricals: 
# Personal.Loan [target outcome], Securities.Account, CD.Account, Online,
# CreditCard, Education [this needs to be split]
# Zip Code and ID (arith not going to do anything here) but we're dropping

# Numeric:
# Family (even though count is small), Age, Experience, CCAvg, Incomem, Mortgage

```
### Bad data || but instructions say to keep all predictors except-listed.
It was seen.
```{r}
# summary showed -3 experience value;... assuming this is correlated to -3 years
# it's likely bad data -- these rows //should// be removed.
# especially since all - experience customers declined the loan.
summary(uBank$Experience)
sort(uBank$Experience)
uBank[uBank$Experience < 0, ]
uBank[uBank$Experience < 0, c("Experience", "Personal.Loan")]
# verify they're all negative. true == all declined.
# this is === if ALL people with negative experience are decliend, print true
all(uBank$Personal.Loan[uBank$Experience<0] == 0)
```


### Customer definition from the assignment (for q1, q4)

```{r}
sample_customer <- data.frame(
  Age = 40,
  Experience = 10,
  Income = 84,
  Family = 2,
  CCAvg = 2,
  Education_1 = 0,
  Education_2 = 1,
  Education_3 = 0,
  Mortgage = 0,
  Securities.Account = 0,
  CD.Account = 0,
  Online = 1,
  CreditCard = 1
)
```

### Sample has Edu1,Edu2, Edu3 so we can factor edu
Set diff shows id / zip / edu / pers loan as differences
once edu is factored we will remove id, zip, personal loan
(though i would have assumed zip would be relevant, location matters).
It would have been cool to geoplot them on a heatmap.

```{r}
setdiff(names(uBank), names(sample_customer))
```
```{r}

```

