---
title: "Assignment_2"
author: "Joseph Demore"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# short block to check if the packages are installed; will download and install any that are not.
packages <- c("class", "caret", "gmodels")

for (pkg in packages) {
  if (!requireNamespace(pkg, quietly = TRUE)) install.packages(pkg, repos = "https://cloud.r-project.org")
  library(pkg, character.only = TRUE)
}
```

```{r}
# helper fn's

# fn to count unique
uCount <- function(x) {
  length(unique(x))
}

# normalizations for data
normalize <- function(data, model) {
  data[qVars] <- predict(model, data[qVars])
  data[predictors]
}

```

### Data Setup & Inspection
Sorted Ascending by unique values to help id cat's. not dogs. cats.

```{r}
# Reads UniversalBank.csv; str shows structure (colnam, type, 10 values)
uBank <- read.csv("UniversalBank.csv")
str(uBank)

# summary to check data
summary(uBank)

# uCount on each col to help identify categoricals
sapply(uBank, uCount)

# sorted
sort(sapply(uBank, uCount))

# categoricals: 
# Personal.Loan [target outcome], Securities.Account, CD.Account, Online,
# CreditCard, Education [this needs to be split]
# Zip Code and ID (arith not going to do anything here) but we're dropping

# Numeric:
# Family (even though count is small), Age, Experience, CCAvg, Incomem, Mortgage

```
### Bad data || but instructions say to keep all predictors except-listed.
It was seen.

```{r}
# summary showed -3 experience value;... assuming this is correlated to -3 years
# it's likely bad data -- these rows //should// be removed.
# especially since all - experience customers declined the loan.
summary(uBank$Experience)
sort(uBank$Experience)
uBank[uBank$Experience < 0, ]
uBank[uBank$Experience < 0, c("Experience", "Personal.Loan")]
# verify they're all negative. true == all declined.
# this is === if ALL people with negative experience are decliend, print true
all(uBank$Personal.Loan[uBank$Experience<0] == 0)
```


### Customer definition from the assignment (for q1, q4)

```{r}
sCust <- data.frame(
  Age = 40,
  Experience = 10,
  Income = 84,
  Family = 2,
  CCAvg = 2,
  Education_1 = 0,
  Education_2 = 1,
  Education_3 = 0,
  Mortgage = 0,
  Securities.Account = 0,
  CD.Account = 0,
  Online = 1,
  CreditCard = 1
)
```

### Sample has Edu1,Edu2, Edu3 so we can factor edu
Set diff shows id / zip / edu / pers loan as differences
once edu is factored we will remove id, zip, personal loan
(though i would have assumed zip would be relevant, location matters).
It would have been cool to geoplot them on a heatmap.

```{r}
setdiff(names(uBank), names(sCust))
```
### Drop ID, Zip then split nonbinary categoricals

```{r}
# categorizes education via factor, binds to df and removes dropped cols
# note sep by _ is only there to match the sample customer.
uBank$Education <- factor(uBank$Education)
uBank <- cbind(uBank, predict(dummyVars(~ Education, data = uBank, sep = "_"), uBank))
uBank$Education <- NULL
uBank$ID <- NULL
uBank$ZIP.Code <- NULL

summary(uBank)
```

```{r}
# only remaining Personal.Loan (acceptance) is missing from the sample customer.
setdiff(names(uBank), names(sCust))
```
### Let's do some ML.. finally.

```{r}
# this one needs to be categorical since it's our target. other cat binarys need
# to stay 0-1 for euclid distance on knn.
uBank$Personal.Loan <- factor(uBank$Personal.Loan)

# make our randomness not randomness at all.
set.seed(1234567890)
# partition for t&v
split = createDataPartition(uBank$Personal.Loan, p=0.6, list = FALSE)
# training gets 60% and validation 40% // split indexes at 60% of data
training <- uBank[split,]
validation <- uBank[-split,]
# define predictors; which are all cols except personal loan //via setdiff
predictors <- setdiff(names(uBank), "Personal.Loan")
# define quant & cat: setdiff predictors&qVars to get cats. === (1- X) for names
qVars <- c("Age", "Experience", "Income", "Family", "CCAvg", "Mortgage")
cVars <- setdiff(predictors, qVars)
```

### Normalization...
Note I defined this above myself.

```{r}
# normalization of model, training, validation, and sample customer
normModel <- preProcess(training[qVars])
tNorm <- normalize(training, normModel)
vNorm <- normalize(validation, normModel)
cNorm <- normalize(sCust, normModel)
```

